---
title: "NBA-analytics"
author: "G"
date: "2022-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# init
```{r echo=FALSE,results='hide'}
#initialize packages
library(GGally)
library(ggplot2)
library(moments)
```

To check missing cases in dataframe:

which(! complete.cases("dataframe"))

#```{r}
reduced_data = read.csv("reduced_data_odd.csv")
rownames(reduced_data) = reduced_data$X
reduced_data$X = NULL
head(reduced_data)
#```


```{r}
cumulative_data = read.csv("cumulative_processed_data_22.csv")
rownames(cumulative_data) = cumulative_data$X

full_data = data.frame(cumulative_data)

cumulative_data$X = NULL
cumulative_data$X3Pcent = NULL
cumulative_data$FTcent = NULL
cumulative_data$X2Pcent = NULL
cumulative_data$FGcent = NULL
head(cumulative_data)
```
# profile data

## curry distribution
```{r}
players = unique(cumulative_data$Player)
length(players)
```
```{r}
#curry.data = cumulative_data[cumulative_data$Player == "Stephen Curry",]

for(i in 1:20){
  cat("\nPlotting data for:", players[i])
  playerData = cumulative_data[cumulative_data$Player == players[i],]
  g = ggplot()+geom_histogram(data = playerData,aes(x=X2P))
  print(g)
}

```


## distributions
```{r}
#Date,Age,Game Result,GS,MP,FG,FGA,FGcent,3P,3PA,3Pcent,FT,FTA,FTcent,ORB,DRB,TRB,AST,STL,BLK,TOV,PF,PTS,GmSc,PointDifference,Player,2P,2PA,2Pcent
ggplot(data=cumulative_data,aes(x=Age))+geom_histogram()
ggplot(data=cumulative_data,aes(x=MP))+geom_histogram(binwidth=1)
X2P_P=ggplot(data=cumulative_data,aes(x=X2P))+geom_histogram(binwidth=1)
print(X2P_P)
X2PA_P=ggplot(data=cumulative_data,aes(x=X2PA))+geom_histogram(binwidth=1)
print(X2PA_P)
X3P_P =ggplot(data=cumulative_data,aes(x=X3P))+geom_histogram(binwidth=1)
print(X3P_P)
X3PA_P = ggplot(data=cumulative_data,aes(x=X3PA))+geom_histogram(binwidth=1)
print(X3PA_P)
#ggplot(data=cumulative_data,aes(x=))+geom_histogram()
```

```{r}
sd(cumulative_data$X2P)

```

```{r}
p = ggplot_build(X2PA_P)
p$data
```

sp=split(p$data[c("x", "density")],p$data[1])
new_weight <- 2
sapply(sp, \(DF){
  with(DF, approx(x, density, xout = new_weight))
})


```{r}

#Date,Age,Game Result,GS,MP,FG,FGA,FGcent,3P,3PA,3Pcent,FT,FTA,FTcent,ORB,DRB,TRB,AST,STL,BLK,TOV,PF,PTS,GmSc,PointDifference,Player,2P,2PA,2Pcent
ggplot(data=cumulative_data,aes(x=Age))+geom_density()
ggplot(data=cumulative_data,aes(x=MP))+geom_density()
ggplot(data=cumulative_data,aes(x=FG))+geom_density()
ggplot(data=cumulative_data,aes(x=FGA))+geom_density()
ggplot(data=cumulative_data,aes(x=X3P))+geom_density()
ggplot(data=cumulative_data,aes(x=X3PA))+geom_density()
ggplot(data=cumulative_data,aes(x=BLK))+geom_histogram()
ggplot(data=cumulative_data,aes(x=PointDifference))+geom_density()
#ggplot(data=cumulative_data,aes(x=))+geom_histogram()
ggplot(data=cumulative_data,aes(x=PTS))+geom_density()
```


```{r}
mean(cumulative_data$X3PA)

qpois(.95,5.157895)
```



## Profiling (skewness/kurtosis)
```{r}

skewness(cumulative_data$FG)
kurtosis(cumulative_data$FG)
```


## our calculated league means
```{r}
mean(cumulative_data$X2P)

mean(cumulative_data$X2PA)
```

```{r}
mean(cumulative_data$X3P)
mean(cumulative_data$X3PA)
```


## find datas
```{r}
#1685 3426

full_data[1457,]
#min(full_data['PointDifference'])
#max(full_data['PointDifference'])

```

```{r}
which(! complete.cases(cumulative_data))
which(full_data['PointDifference']== 10)
```


```
dataForGGpairs = data.frame(cumulative_data)
dataForGGpairs$Date = NULL
dataForGGpairs$Player = NULL
head(dataForGGpairs)
#ggpairs(data=dataForGGpairs,lower = list(continuous = "smooth_loess", combo = "facethist", discrete = "facetbar", na = "na"))
```

# Plot to MP

```{r}
ggplot(data=cumulative_data,aes(x=MP,y=PTS))+geom_point(alpha=0.1)+geom_smooth(color="red")
```


# Base Model

```{r}
offensive.model1 = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV+PF),data=cumulative_data)
summary(offensive.model1)
```
```{r}
simpleModel = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV),data=cumulative_data)
summary(simpleModel)
```
### check for non-linear
Check for non-linear relationship between X and PointDifference
```{r}
#Age+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+MP+STL+AST+BLK+TOV+PF
ggplot(data=cumulative_data,aes(x=Age,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=MP,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=X3P,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=X3PA,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=X2P,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=X2PA,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=ORB,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=DRB,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=FT,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=FTA,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=STL,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=AST,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=BLK,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=TOV,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=PF,y=PointDifference))+geom_point()
ggplot(data=cumulative_data,aes(x=,y=PointDifference))+geom_point()
```
```{r}
ggplot(data=full_data,aes(x=FTcent,y=PointDifference))+geom_point()+geom_smooth()
ggplot(data=full_data,aes(x=FGcent,y=PointDifference))+geom_point()+geom_smooth()
ggplot(data=full_data,aes(x=X2Pcent,y=PointDifference))+geom_point()+geom_smooth()
ggplot(data=full_data,aes(x=X3Pcent,y=PointDifference))+geom_point()+geom_smooth()
```



```{r}
offensive.model1.interact = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+MP+STL+AST+BLK+TOV)^2,data=cumulative_data)
summary(offensive.model1.interact)
```

```{r}
#using p = 0.05
model.interaction.reduced = lm(PointDifference~(factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+MP+STL+AST+BLK+TOV+ORB*AST+X3PA*BLK+MP*BLK+MP*AST+MP*X2P+MP*X2PA+factor(GS)*DRB+factor(GS)*BLK),data=cumulative_data)
summary(model.interaction.reduced)
```
```{r}
#using p = 0.05
model.interaction.reduced_further = lm(PointDifference~(factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+MP+STL+AST+TOV+MP*AST+MP*X2P+MP*X2PA+factor(GS)*DRB),data=cumulative_data)
summary(model.interaction.reduced_further)
```

## Prediction
```{r}
#simpleModel = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV),data=cumulative_data)
player1.a = data.frame(Age=23,GS=1,MP=36.66,X3P=2,X3PA=10,X2P=8,X2PA=13,FT=13,FTA=13,ORB=0,DRB=9,STL=1,AST=6,BLK=0,TOV=4)
player1.b = data.frame(Age=23,GS=1,MP=29.95,X3P=4,X3PA=9,X2P=8,X2PA=14,FT=4,FTA=5,ORB=2,DRB=5,STL=2,AST=10,BLK=2,TOV=3)
player2.a = data.frame()
predict(model.interaction.reduced_further,player1.a,interval = 'predict')
predict(simpleModel,player1.a,interval ='predict')
```
### -40 game
```{r}
predict(model.interaction.reduced_further,full_data[831,],interval = 'predict')
predict(simpleModel,full_data[831,],interval ='predict')
```

# Positive Point Prediction
```{r}
positivePoints = read.csv("positive_points.csv")
simpleModel.pp = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV),data=positivePoints)
summary(simpleModel.pp)
```
### reduced simple model

```{r}
reduced.pp = lm(PointDifference~(MP+X3P+X3PA+X2P+X2PA+DRB+STL+AST+TOV),data=positivePoints)
summary(reduced.pp)
```
### check for non-linear
Check for non-linear relationship between X and PointDifference
```{r}
#Age+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+MP+STL+AST+BLK+TOV+PF
ggplot(data=positivePoints,aes(x=Age,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=MP,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=X3P,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=X3PA,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=X2P,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=X2PA,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=ORB,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=DRB,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=FT,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=FTA,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=STL,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=AST,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=BLK,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=TOV,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
ggplot(data=positivePoints,aes(x=PF,y=PointDifference))+geom_point(alpha=.1)+geom_smooth()
#ggplot(data=positivePoints,aes(x=,y=PointDifference))+geom_point()
```

### interact
```{r}
interaction.pp = lm(PointDifference~(MP+X3P+X3PA+X2P+X2PA+DRB+STL+AST+TOV)^2,data=positivePoints)
summary(interaction.pp)
```
### reduced interact
```{r}
reduced.interact.pp = lm(PointDifference~(MP+X3P+X2P+DRB+STL+AST+TOV+DRB*AST+X2P*STL+X3P*TOV+X3P*DRB+MP*AST+MP*DRB),data=positivePoints)
summary(reduced.interact.pp)
```

#### refinement
```{r}
reduced.interact.pp = lm(PointDifference~(MP+X3P+X2P+DRB+AST+TOV+DRB*AST+X3P*TOV+MP*AST+MP*DRB),data=positivePoints)
summary(reduced.interact.pp)
```


```{r}
interaction.pp = lm(PointDifference~(MP+X3P+X3PA+X2P+X2PA+DRB+STL+AST+TOV)^2,data=positivePoints)
summary(interaction.pp)
```




# Negative Point Prediction
```{r}
negativePoints = read.csv("negative_points.csv")
simpleModel.np = lm(PointDifference~(Age+factor(GS)+MP+X3P+X3PA+X2P+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV),data=negativePoints)
summary(simpleModel.np)
```

### reduced simple model
```{r}
reduced.np = lm(PointDifference~(factor(GS)+MP+X3P+X3PA+X2P+X2PA+DRB+AST+BLK+TOV),data=negativePoints)
summary(reduced.np)
```
### interact
```{r}
interact.np = lm(PointDifference~(factor(GS)+MP+X3P+X3PA+X2P+X2PA+DRB+AST+BLK+TOV)^2,data=negativePoints)
summary(interact.np)
```

### reduced interact
```{r}
reduced.interact.np = lm(PointDifference~GS+MP+X3PA+X2P+X2PA+DRB+AST+BLK+TOV+factor(GS):MP+factor(GS):X2PA+factor(GS):DRB+MP:X2PA+MP:DRB+X3P:AST+X3PA:AST+X3PA:BLK+X2PA:TOV,data=negativePoints)
summary(reduced.interact.np)
```
#### refinement
```{r}
reduced.interact.np = lm(PointDifference~GS+MP+X3PA+X3P+X2P+X2PA+DRB+AST+BLK+TOV+factor(GS):MP+factor(GS):X2PA+factor(GS):DRB+MP:X2PA+X3PA:BLK,data=negativePoints)
summary(reduced.interact.np)
```




# Testing
### -40 Game
```{r}
predict(reduced.pp,full_data[831,],interval = 'predict')
predict(reduced.pp,full_data[831,],interval ='predict')

```
### +10 game
```{r}
which(full_data['PointDifference']== 10)
#36   64   73   90   93  103  152  292  309  312  345  376  377  385  465  495  578 650  744  745  755  767  828  878  897  971 1019 1021 1075 1080 1107 1120 1190 1199
# 1367 1373 1389 1409 1447 1457 1461 1524 1543 1624 1625 1748 1810 2118 2175 2184 2212
# 2233 2238 2246 2319 2435 2447 2466 2475 2482 2485 2500 2555 2578 2584 2600 2663 2693

predict(reduced.pp,full_data[1457,],interval='predict')
```
### +15 game
```{r}
test = predict(reduced.pp,full_data[i,],interval='predict')
print(data.frame(test)$fit)
```



### +40game
```{r}
#1685 3426
which(full_data['PointDifference']== 40)
predict(reduced.pp,full_data[1685,],interval='predict')
```

### iterative testing positive

```{r}
simulatePP = function(x){
gamesThatMatch = which(full_data['PointDifference']== x)
simulatedValues = numeric(length(gamesThatMatch))
for (i in 1:length(gamesThatMatch)){
  interval = predict(reduced.interact.pp,full_data[gamesThatMatch[i],],interval='predict')
  simulatedValues[i] = data.frame(interval)$fit
}
res= mean(simulatedValues)
return(res)
}
pos.x=(0:44)
#print(x)
pos.y = numeric(length(pos.x))
for (i in 1:length(pos.x)){
  pos.y[i] = simulatePP(pos.x[i])
}
#print(y)
pos.testFrame = data.frame(x=pos.x,y=pos.y)
head(pos.testFrame)
```
```{r}
ggplot(data=pos.testFrame,aes(x=x,y=y))+geom_point()
```

### iterative testing negative

```{r}
simulateNP = function(x){
gamesThatMatch = which(full_data['PointDifference']== x)
simulatedValues = numeric(length(gamesThatMatch))
for (i in 1:length(gamesThatMatch)){
  interval = predict(reduced.interact.np,full_data[gamesThatMatch[i],],interval='predict')
  simulatedValues[i] = data.frame(interval)$fit
}
res= mean(simulatedValues)
return(res)
}
neg.x=(-40:0)
#print(x)
neg.y = numeric(length(neg.x))
for (i in 1:length(neg.x)){
  neg.y[i] = simulateNP(neg.x[i])
}
#print(y)
neg.testFrame = data.frame(x =neg.x,y= neg.y)
head(neg.testFrame)
```

#### ALL
```{r}

simulatedValues = numeric(nrow(positivePoints))
xvals = numeric(nrow(positivePoints))

for (i in 1:nrow(positivePoints)){
  interval = predict(reduced.interact.pp,positivePoints[i,],interval='predict')
  simulatedValues[i] = data.frame(interval)$fit
  xvals[i] = positivePoints[i,]$PointDifference
}
print(length(positivePoints))
print(simulatedValues)
fullSimulation.pp = data.frame(PointDifference=xvals,simulatedValues)
head(fullSimulation.pp)
```
```{r}
ggplot(data=fullSimulation.pp,aes(x=PointDifference,y=simulatedValues))+geom_point(alpha=0.1,color="blue")+geom_smooth()
```



```{r}
ggplot(data=neg.testFrame,aes(x=x,y=y))+geom_point()
```

### combined plot
```{r}
combined = rbind(pos.testFrame,neg.testFrame)
ggplot(data=combined,aes(x=x,y=y))+geom_point()
```

# Point Prediction Model

```{r}
pointModel = lm(X2P~(Age+factor(GS)+MP+X3P+X3PA+X2PA+ORB+DRB+FT+FTA+STL+AST+BLK+TOV+PF),data=cumulative_data)
summary(pointModel)
```
## Reduced
```{r}
pointModel.reduced = lm(X2P~(Age+X3PA+X2PA+ORB+DRB+FTA+AST+BLK+TOV+PF),data=cumulative_data)
summary(pointModel.reduced)
```
## Interactions
```{r}
pointModel.interact = lm(X2P~(Age+X3PA+X2PA+ORB+DRB+FTA+AST+BLK+TOV+PF)^2,data=cumulative_data)
summary(pointModel.interact)
```
## reduced interact

```{r}
pointModel.interact.reduced = lm(X2P~(Age+X3PA+X2PA+ORB+DRB+FTA+AST+BLK+TOV+PF+AST*TOV+AST*BLK+FTA*TOV+FTA*BLK+ORB*TOV+ORB*FTA+X2PA*AST+X2PA*FTA+X2PA*DRB+X3PA*TOV+X3PA*ORB+Age*X2PA),data=cumulative_data)
summary(pointModel.interact.reduced)
```


